---
title: "Getting Started with bsvarPost"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with bsvarPost}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
<!-- generated-by: gsd-doc-writer -->

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>",
                      fig.width = 7, fig.height = 4)
library(bsvarPost)
post <- readRDS("fixtures/fiscal_post_bsvar.rds")
post_alt <- readRDS("fixtures/fiscal_post_bsvar_alt.rds")
```

`bsvarPost` starts where `bsvars` and `bsvarSIGNs` stop: with an estimated
posterior. It turns posterior results into tidy tables, cumulative responses,
comparisons, probability statements, coherent representative draws, event
summaries, and publication-ready output.

This guide assumes that you already know how to specify and estimate a model
with one of the parent packages. In one short workflow, we examine the response
of US GDP (`gdp`) to the government-spending shock (`gs`) in the
`bsvars::us_fiscal_lsuw` data. The included fixtures keep the article fast; use
the same calls with your own posterior object.

## Start from a posterior

Only the estimation handoff matters here. For example, a minimal `bsvars`
workflow is:

```{r estimate, eval = FALSE}
data("us_fiscal_lsuw", package = "bsvars")
spec <- bsvars::specify_bsvar$new(us_fiscal_lsuw, p = 1)
post <- bsvars::estimate(spec, S = 2000, thin = 1,
                         show_progress = FALSE)
```

The rest of the guide uses the precomputed `post` loaded above. It is a regular
`PosteriorBSVAR` object, so there is no import or conversion step.

## Use one tidy extraction pattern

The canonical pattern is: pass the posterior to a `tidy_*()` function, then
filter the resulting tibble to the research question. Here we request 90%
pointwise intervals and keep four horizons from the GDP response to a spending
shock.

```{r tidy-irf}
irf_tbl <- tidy_irf(post, horizon = 12, probability = 0.90)
irf_focus <- subset(irf_tbl, variable == "gdp" & shock == "gs" &
                    horizon %in% c(0, 4, 8, 12),
                    select = c(variable, shock, horizon, median, lower, upper))
irf_focus
```

The same shape extends to other posterior quantities. Summary tables contain
posterior location and interval columns; set `draws = TRUE` when downstream
work needs individual draws.

```{r tidy-family, eval = FALSE}
tidy_cdm(post, horizon = 12)       # cumulative responses
tidy_fevd(post, horizon = 12)      # variance shares
tidy_hd(post)                      # historical contributions
tidy_forecast(post, horizon = 8)   # predictive paths
tidy_shocks(post)                  # structural shocks
```

This is usually enough to join results to annotations, filter panels, or use a
different graphics system without unpacking posterior arrays by hand.

## Accumulate responses with `cdm()`

`cdm()` retains the posterior draws while accumulating the response at each
horizon. `tidy_cdm()` then applies the same extraction pattern used above.

```{r cdm}
cdm_draws <- cdm(post, horizon = 12)
cdm_tbl <- tidy_cdm(cdm_draws, probability = 0.90)
cdm_focus <- subset(cdm_tbl, variable == "gdp" & shock == "gs" &
                    horizon %in% c(0, 4, 8, 12),
                    select = c(variable, shock, horizon, median, lower, upper))
cdm_focus
```

These are cumulative dynamic responses. Whether they have an economic
multiplier interpretation depends on the model's transformations and scaling;
see `?cdm` for the `scale_by` and `scale_var` options.

## Compare specifications in the same table

The second fixture is the same three-variable system with a longer lag order.
Named posterior arguments become stable labels in the `model` column.

```{r compare}
cdm_comparison <- compare_cdm(
  baseline = post,
  longer_lag = post_alt,
  horizon = 12,
  probability = 0.90
)
subset(cdm_comparison, variable == "gdp" & shock == "gs" &
       horizon %in% c(4, 8, 12),
       select = c(model, horizon, median, lower, upper))
```

Use the same pattern with `compare_irf()`, `compare_fevd()`, and the other
`compare_*()` helpers. The comparison is descriptive: it puts like-for-like
posterior summaries together but does not choose a preferred specification.

## Ask a posterior probability question

Suppose the question is: *what is the posterior probability that the GDP
response to the spending shock is positive at each selected horizon?*

```{r hypothesis}
prob_positive <- hypothesis_irf(
  post,
  variables = "gdp",
  shocks = "gs",
  horizon = c(0, 4, 8, 12),
  relation = ">",
  value = 0
)
prob_positive[, c("variable", "shock", "horizon", "posterior_prob")]
```

`posterior_prob` is a model-conditional posterior probability, not a
frequentist p-value. For questions that must hold across several horizons use
`joint_hypothesis_irf()`; for economically meaningful thresholds use
`magnitude_audit()`; and for a band covering an entire selected path use
`simultaneous_irf()`. The
[Inference and Comparison](inference-and-comparison.html) article develops
these distinctions.

## Keep one coherent draw and summarize timing

Pointwise medians need not correspond to any single posterior draw.
`median_target_irf()` instead selects one internally coherent draw closest to
the median response path over the requested subset.

```{r representative}
representative <- median_target_irf(
  post,
  horizon = 12,
  variables = "gdp",
  shocks = "gs"
)
representative_path <- subset(summary(representative),
                              variable == "gdp" & shock == "gs")
subset(representative_path, horizon %in% c(0, 4, 8, 12),
       select = c(variable, shock, horizon, median, draw_index, method))
```

For a compact timing statement, `peak_response()` reports both the response
level and the horizon of the peak for every posterior draw, then summarizes
those quantities.

```{r timing}
peak <- peak_response(
  post,
  horizon = 12,
  variables = "gdp",
  shocks = "gs",
  absolute = TRUE
)
peak[, c("variable", "shock", "median_value", "median_horizon")]
```

For persistence questions, `duration_response()`, `half_life_response()`, and
`time_to_threshold()` provide related summaries without reducing the posterior
to a single curve.

## Move from an HD path to an event question

Start with draw-level `tidy_hd()` data, then choose a window
that is substantively meaningful for the application. This illustrative
four-quarter window shows the handoff from event aggregation to shock ranking.

```{r hd-event}
hd_draws <- tidy_hd(post, draws = TRUE)
event_tbl <- tidy_hd_event(
  hd_draws,
  start = "1958",
  end = "1958.75"
)

subset(event_tbl, variable == "gdp",
       select = c(variable, shock, event_start, event_end, median))

ranking <- shock_ranking(hd_draws, start = "1958", end = "1958.75",
                         variables = "gdp")
ranking[, c("variable", "shock", "median", "rank")]
```

This adds event-level questions to the parent package's historical
decomposition rather than replacing its standard HD workflow. Window choice
and contribution interpretation are covered in
[Historical-Decomposition Events](historical-decomposition-events.html).

## Finish with publication-ready output

One call turns the representative draw into a consistently styled `ggplot2`
figure. The returned object remains a normal `ggplot`, so annotations can still
be added before saving.

```{r publication-plot}
publish_bsvar_plot(representative_path, family = "irf", preset = "paper",
  title = "GDP response to a government-spending shock",
  caption = "Median-target posterior draw")
```

For a table, keep one backend and the same focused result:

```{r publication-table}
as_kable(peak, caption = "Peak absolute GDP response",
         digits = 3, preset = "compact")
```

## Where next?

- [Inference and Comparison](inference-and-comparison.html) covers joint and
  magnitude questions, simultaneous bands, and richer model comparisons.
- [Historical-Decomposition Events](historical-decomposition-events.html)
  covers event windows, contribution views, and shock ranking.
- [Sign-Restricted Workflows](sign-restricted-workflows.html) covers
  `most_likely_admissible_*()`, `restriction_audit()`,
  `acceptance_diagnostics()`, and sign-model comparisons.

The [reference index](https://davidzenz.github.io/bsvarPost/reference/index.html) lists variants and lower-level
helpers when a workflow needs more control.
