---
title: "Getting Started with bsvarPost"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with bsvarPost}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

`bsvarPost` is a companion post-estimation package for `bsvars` and
`bsvarSIGNs`. It does not estimate models itself. Instead, it adds a consistent
layer for:

- cumulative dynamic multipliers via `cdm()`
- tidy extraction helpers
- model comparison utilities
- `ggplot2` plotting methods
- bridges to downstream workflows such as `tsibble` and `APRScenario`
- representative-model summaries and posterior audit helpers

## Example: `bsvars`

```{r}
library(bsvars)
library(bsvarPost)

data(us_fiscal_lsuw)
set.seed(1)

spec <- specify_bsvar$new(us_fiscal_lsuw, p = 1)
post <- estimate(spec, S = 100, thin = 1, show_progress = FALSE)
```

Compute cumulative dynamic multipliers:

```{r}
cdm_obj <- cdm(post, horizon = 8)
summary(cdm_obj)
plot(cdm_obj)
```

Normalize by the sample standard deviation of the shocked variable only if that
is the reporting convention you want:

```{r}
cdm_scaled <- cdm(post, horizon = 8, scale_by = "shock_sd")
```

Extract tidy tables for downstream analysis:

```{r}
irf_tbl <- tidy_irf(post, horizon = 8)
cdm_tbl <- tidy_cdm(post, horizon = 8)
fevd_tbl <- tidy_fevd(post, horizon = 8)
fc_tbl <- tidy_forecast(post, horizon = 8)
```

Plot tidy outputs with `ggplot2` defaults:

```{r}
ggplot2::autoplot(irf_tbl)
ggplot2::autoplot(cdm_tbl)
```

Representative-model summaries are useful when you want a single draw
that is close to the posterior center rather than pointwise medians that need
not correspond to one structural model.

```{r}
rep_irf <- median_target_irf(post, horizon = 8)
summary(rep_irf)
plot(rep_irf)
```

Posterior hypothesis and magnitude statements can be evaluated directly from the
response draws.

```{r}
hypothesis_irf(post, variable = 1, shock = 1, horizon = 4, relation = ">", value = 0)
magnitude_audit(post, type = "cdm", variable = 1, shock = 1, horizon = 8, relation = ">", value = 0)
```

## Example: `bsvarSIGNs`

`bsvarPost` uses the same post-estimation interface for sign-restricted models.

```{r}
library(bsvarSIGNs)
library(bsvarPost)

data(optimism)

sign_irf <- matrix(c(0, 1, rep(NA, 23)), 5, 5)

spec <- specify_bsvarSIGN$new(
  optimism * 100,
  p = 4,
  sign_irf = sign_irf
)

post <- estimate(spec, S = 100, thin = 1, show_progress = FALSE)
```

Compute and summarize CDMs:

```{r}
cdm_obj <- cdm(post, horizon = 12)
summary(cdm_obj)
plot(cdm_obj)
```

Extract tidy summaries:

```{r}
irf_tbl <- tidy_irf(post, horizon = 12)
cdm_tbl <- tidy_cdm(post, horizon = 12)
fevd_tbl <- tidy_fevd(post, horizon = 12)
```

`bsvarPost` also supports representative sign-restricted summaries and
restriction audits. `median_target_*()` selects the stored draw closest to the
posterior center. `most_likely_admissible_*()` ranks stored admissible draws by
the reconstructed sign-restriction score available from `bsvarSIGNs`.

```{r}
rep_median <- median_target_irf(post, horizon = 12)
rep_mla <- most_likely_admissible_irf(post, horizon = 12)
summary(rep_median)
summary(rep_mla)

audit_tbl <- restriction_audit(post)
audit_tbl
```

Magnitude restrictions in `bsvarPost` are post-estimation summaries, not
identification constraints.

```{r}
magnitude_audit(post, type = "irf", variable = 2, shock = 1, horizon = 0, relation = ">", value = 0)
```

## Comparing models

Comparison helpers accept several posterior objects and return tidy comparison
tables.

```{r}
cmp_irf <- compare_irf(
  p1 = post,
  p2 = post,
  horizon = 8
)

cmp_cdm <- compare_cdm(
  p1 = post,
  p2 = post,
  horizon = 8
)

ggplot2::autoplot(cmp_irf)
ggplot2::autoplot(cmp_cdm)
```

## `tsibble` bridge

If `tsibble` is installed, tidy outputs can be converted directly:

```{r}
library(tsibble)

fc_tbl <- tidy_forecast(post, horizon = 8)
fc_tsbl <- as_tsibble_post(fc_tbl)
```

For IRF, CDM, and FEVD outputs, `horizon` is used as the index. For forecast,
shock, and historical decomposition outputs, `time` is used when available.

## APRScenario bridge

`APRScenario` uses forecast summaries with columns `hor`, `variable`, `lower`,
`center`, and `upper`. `bsvarPost` provides converters in both directions.

```{r}
fc_tbl <- tidy_forecast(post, horizon = 8)

apr_tbl <- as_apr_cond_forc(
  fc_tbl,
  origin = as.Date("2024-01-01"),
  frequency = "quarter"
)

back_to_tidy <- tidy_apr_forecast(apr_tbl)
```

If `APRScenario` is installed, you can also forward to its matrix generator:

```{r}
mats <- apr_gen_mats(post, specification = spec)
```
