---
title: "Historical-Decomposition Event Workflows"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Historical-Decomposition Event Workflows}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
<!-- generated-by: gsd-doc-writer -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4
)
library(bsvarPost)
post <- readRDS("fixtures/fiscal_post_bsvar.rds")
post_alt <- readRDS("fixtures/fiscal_post_bsvar_alt.rds")
```

Historical decompositions from `bsvars` and `bsvarSIGNs` describe how structural
shock contributions evolve through the sample. `bsvarPost` adds a narrower
workflow: aggregate those contributions over a substantively chosen window,
rank the shocks within it, and compare the same event across specifications.

This article assumes that you already have a fitted posterior and understand
the identification behind its shock labels. See the parent-package
documentation for estimation and the standard historical-decomposition
workflow. The precomputed fiscal posteriors used here avoid re-estimation and
keep the vignette reproducible.

## Start from draw-level contributions

Event summaries need the posterior draws, not a table that has already collapsed
them to pointwise intervals. The canonical handoff is therefore
[`tidy_hd(draws = TRUE)`](../reference/tidy_hd.html):

```{r hd-draws}
hd_draws <- tidy_hd(post, draws = TRUE)

data.frame(
  first = head(unique(as.character(hd_draws$time)), 1),
  last = tail(unique(as.character(hd_draws$time)), 1),
  draws = length(unique(hd_draws$draw))
)
```

Each row now identifies a model, variable, shock, observation, and posterior
draw. Keep this draw-level table when defining events so uncertainty is
aggregated *after* contributions are summed within the window.

## Choose the window before ranking shocks

An event window is an analytical decision, not something the ranking algorithm
discovers. A defensible workflow usually does three things:

1. Define the dates from the research question or an external chronology,
   before inspecting which shock ranks first.
2. Match `start` and `end` to the model's time labels and use a window consistent
   with the data frequency.
3. Repeat the analysis over reasonable neighboring windows as a sensitivity
   check, without silently replacing the prespecified result.

For this compact example, we use the complete four-quarter span from `1958` to
`1958.75`. It is a convenient, bounded window inside both fixtures—not a claim
that these documentation fixtures establish a particular historical narrative.
They contain only 200 stored draws, and the shock names inherit whatever
interpretation the fitted model's identification supports.

```{r event-window}
event_start <- "1958"
event_end <- "1958.75"

event <- tidy_hd_event(
  hd_draws,
  start = event_start,
  end = event_end,
  probability = 0.90
)

subset(
  event,
  variable == "gdp",
  select = c(variable, shock, event_start, event_end,
             median, lower, upper)
)
```

[`tidy_hd_event()`](../reference/tidy_hd_event.html) sums each shock's
contribution across the inclusive window separately for every draw, then reports
posterior summaries. Set `draws = TRUE` when a downstream calculation needs the
aggregated draws themselves.

Do not read the largest median as certainty about a shock's historical role.
The interval columns retain posterior uncertainty, while the structural meaning
of each label still comes from the original model.

## Rank the contributions relevant to the question

[`shock_ranking()`](../reference/shock_ranking.html) orders shocks separately
within each model-variable panel. Restricting to `gdp` keeps the result aligned
with one research question.

```{r rank-shocks}
ranking <- shock_ranking(
  hd_draws,
  start = event_start,
  end = event_end,
  variables = "gdp",
  ranking = "absolute",
  probability = 0.90
)

ranking[, c("variable", "shock", "median", "lower", "upper", "rank")]
```

`ranking = "absolute"` answers which shock has the largest median contribution
in magnitude. Use `ranking = "signed"` only when the ordering by direction is
the intended estimand. In either case, `rank` summarizes the point estimate; it
does not turn overlapping posterior intervals into a decisive ordering.

One useful event figure is the ranked contribution plot. It keeps the sign and
scale of the contribution visible while ordering the bars by importance.

```{r ranking-plot}
ranking_plot <- publish_bsvar_plot(
  ranking,
  family = "shock_ranking",
  preset = "paper",
  title = "GDP contributions during the selected window",
  caption = "Bars are posterior medians; uncertainty remains in the event table."
)
ranking_plot
```

For a direct contribution plot rather than a ranking, use
[`plot_hd_event()`](../reference/plot_hd_event.html). The more specialized
[`plot_hd_event_share()`](../reference/plot_hd_event_share.html),
[`plot_hd_event_cumulative()`](../reference/plot_hd_event_cumulative.html), and
[`plot_hd_event_distribution()`](../reference/plot_hd_event_distribution.html)
are documented in the reference index; use them when the question genuinely
concerns composition, within-window accumulation, or the full draw distribution.

## Hold the event fixed across specifications

Changing a lag order or prior should not also change the event definition.
[`compare_hd_event()`](../reference/compare_hd_event.html) applies the identical
window to named posterior objects and returns a tidy model column.

```{r compare-event}
event_comparison <- compare_hd_event(
  baseline = post,
  alternative = post_alt,
  start = event_start,
  end = event_end,
  probability = 0.90
)

comparison_gdp <- subset(
  event_comparison,
  variable == "gdp",
  select = c(model, shock, event_start, event_end,
             median, lower, upper)
)
comparison_gdp
```

This comparison is descriptive. Stability across columns can support a
robustness discussion, but `compare_hd_event()` neither selects a preferred
model nor computes a posterior probability that the specifications differ.
When comparing additional windows, preserve names such as `baseline` and
`alternative` so tables and figures remain traceable.

## Hand results to a report

Finish with one focused table and save the already styled plot. Both objects
remain ordinary R outputs, so captions, annotations, and journal-specific
formatting can be added downstream.

```{r publication-table}
as_kable(
  comparison_gdp,
  caption = "GDP shock contributions in the selected event window",
  digits = 2,
  preset = "compact"
)
```

```{r save-plot, eval = FALSE}
ggplot2::ggsave(
  "shock-ranking-1958.pdf",
  ranking_plot,
  width = 7,
  height = 4
)
```

For the broader path from posterior objects to response comparisons and
probability statements, return to [Getting Started](bsvarPost.html). For
sign-restricted identification diagnostics, continue to the dedicated
bsvarSIGNs workflow article.
