---
title: "Inference and Comparison"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Inference and Comparison}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
<!-- generated-by: gsd-doc-writer -->

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(bsvarPost)
post <- readRDS("fixtures/fiscal_post_bsvar.rds")
post_alt <- readRDS("fixtures/fiscal_post_bsvar_alt.rds")
```

This article starts with estimated posteriors and asks questions that cannot be
answered by plotting a pointwise median alone. The running example compares two
lag specifications for the `bsvars::us_fiscal_lsuw` system. We focus on the GDP
response to a government-spending shock, without presuming that its sign or
magnitude is known in advance.

The fixtures contain 200 stored draws and keep the article quick to build. Replace
them with your own compatible `bsvars` or `bsvarSIGNs` posterior objects.

## Is the conclusion sensitive to specification?

Start with the comparison rather than two separate tables. Named arguments to
`compare_irf()` become stable model labels, and every other column has the same
meaning as in `tidy_irf()`.

```{r compare-specifications}
comparison <- compare_irf(
  baseline = post,
  longer_lag = post_alt,
  horizon = 12,
  probability = 0.90
)

subset(
  comparison,
  variable == "gdp" & shock == "gs" & horizon %in% c(0, 4, 8, 12),
  select = c(model, horizon, median, lower, upper))
```

Read this descriptively: shifts in the center or interval show specification
sensitivity; they do not constitute a model-selection rule. The same named-
posterior pattern is available for [CDMs](../reference/compare_cdm.html),
[FEVDs](../reference/compare_fevd.html),
[forecasts](../reference/compare_forecast.html), and
[historical-decomposition events](../reference/compare_hd_event.html).

## How much posterior support does a statement have?

### Pointwise support

For a claim at particular horizons, ask directly what share of posterior draws
satisfies it. Here the claim is that the GDP response is positive.

```{r pointwise-probability}
positive_gdp <- hypothesis_irf(
  post,
  variables = "gdp",
  shocks = "gs",
  horizon = c(0, 4, 8, 12),
  relation = ">",
  value = 0
)

positive_gdp[, c("horizon", "posterior_prob", "median_gap",
                 "lower_gap", "upper_gap")]
```

`posterior_prob` is the model-conditional probability of the stated inequality,
not a p-value. The gap columns summarize the response minus the threshold, so
they retain information about effect size. Use
[`hypothesis_cdm()`](../reference/hypothesis_cdm.html) for the same question
about cumulative responses.

### Joint support across a path

Pointwise probabilities do not answer whether one draw satisfies every
condition. For the stronger claim that the response is positive from impact
through horizon 8, intersect the conditions draw by draw.

```{r joint-probability}
positive_path <- joint_hypothesis_irf(
  post,
  variable = "gdp",
  shock = "gs",
  horizon = 0:8,
  relation = ">",
  value = 0
)

positive_path[, c("posterior_prob", "n_constraints")]
```

The joint probability can be much smaller than the probability at any one
horizon because all nine inequalities must hold in the same draw. See
[`joint_hypothesis_cdm()`](../reference/joint_hypothesis_cdm.html) for cumulative
paths.

### A magnitude question

A sign may be too weak for the research question. Suppose `0.001` is a
pre-specified, substantively meaningful response size on the scale used to fit
the model. We can ask for the probability that the absolute response exceeds
that threshold at horizon 8.

```{r magnitude-question, warning = FALSE}
material_response <- magnitude_audit(
  post,
  type = "irf",
  variable = "gdp",
  shock = "gs",
  horizon = 8,
  relation = ">",
  value = 0.001,
  absolute = TRUE
)

material_response[, c("posterior_prob", "median_gap",
                       "lower_gap", "upper_gap")]
```

Choose the threshold from the variable transformation and application, not
from the posterior output. For cumulative multiplier questions, set
`type = "cdm"` and review the scaling options in
[`magnitude_audit()`](../reference/magnitude_audit.html).

## Does an interval cover the whole path?

Ordinary credible intervals are pointwise. A 90% simultaneous band uses one
sup-norm critical value for the selected response path, giving the stronger
statement that 90% of draws lie inside the band over the selection as a whole.

```{r simultaneous-band}
gdp_band <- simultaneous_irf(
  post,
  horizon = 12,
  probability = 0.90,
  variables = "gdp",
  shocks = "gs"
)

gdp_band[gdp_band$horizon %in% c(0, 4, 8, 12),
         c("horizon", "median", "lower", "upper", "simultaneous_prob")]
```

The band applies only to the variables, shocks, and horizons used to construct
it. Expanding that family generally widens the band. Use
[`simultaneous_cdm()`](../reference/simultaneous_cdm.html) for cumulative
responses and [`plot_simultaneous()`](../reference/plot_simultaneous.html) when
the path is easier to assess graphically.

## Which single draw represents the posterior center?

A grid of pointwise medians need not be generated by any stored parameter draw.
`median_target_irf()` instead selects the one draw closest to the median target
over the requested responses and horizons. Including both GDP and government
spending in the target keeps their displayed paths coherent with each other.

```{r median-target}
representative <- median_target_irf(
  post, horizon = 12,
  variables = c("gdp", "gs"),
  shocks = "gs", horizons = 0:12
)

representative$draw_index

rep_path <- summary(representative)
subset(
  rep_path,
  variable == "gdp" & shock == "gs" & horizon %in% c(0, 4, 8, 12),
  select = c(horizon, median, draw_index, method))
```

Use this draw when a downstream calculation requires one internally consistent
structural model. Do not describe it as posterior uncertainty: retain the full
posterior and its intervals alongside it. The same selection is available for
[CDMs](../reference/median_target_cdm.html); sign-restricted models also have a
distinct [most-likely admissible draw](../reference/most_likely_admissible_irf.html).

## When does the response arrive and fade?

Response-shape summaries operate draw by draw before summarizing, so uncertainty
in timing is not inferred from one median curve. For an interpretable persistence
example, examine the government-spending shock's own response over 20 quarters.

```{r response-timing}
peak <- peak_response(post, horizon = 20, variables = "gs", shocks = "gs",
                      absolute = TRUE)
duration <- duration_response(
  post, horizon = 20, variables = "gs", shocks = "gs",
  relation = ">", value = 0, mode = "consecutive")
half_life <- half_life_response(
  post, horizon = 20, variables = "gs", shocks = "gs", baseline = "peak")
threshold <- time_to_threshold(
  post, horizon = 20, variables = "gs", shocks = "gs",
  relation = "<", value = 0.015)

data.frame(
  question = c("absolute peak", "positive spell", "half-life", "below 0.015"),
  posterior_median = c(
    peak$median_horizon,
    duration$median_duration,
    half_life$median_half_life,
    threshold$median_horizon
  ),
  unit = c("horizon", "periods", "periods after peak", "horizon"),
  reached_prob = c(NA, NA, half_life$reached_prob, threshold$reached_prob)
)
```

`reached_prob` matters for half-life and threshold summaries: a conditional
median can look precise even when few draws reach the event within the requested
horizon. `peak_response()` additionally reports the posterior distribution of
the peak value. Use the dedicated reference pages for alternative definitions:
[`duration_response()`](../reference/duration_response.html),
[`half_life_response()`](../reference/half_life_response.html), and
[`time_to_threshold()`](../reference/time_to_threshold.html).

## Choose the statement that matches the question

- Use a `compare_*()` table to expose specification sensitivity.
- Use `hypothesis_*()` for one or more pointwise posterior probabilities.
- Use `joint_hypothesis_*()` when all selected inequalities must hold together.
- Use `magnitude_audit()` for a pre-specified economically relevant threshold.
- Use `simultaneous_*()` when coverage of an entire selected path is the claim.
- Use a median-target draw only when one coherent stored model is required.
- Use timing summaries for when, how long, and whether a response reaches an
  event—not as substitutes for the response distribution itself.

For sign-restricted posterior diagnostics and admissibility-based representative
draws, continue to [Sign-Restricted Workflows](sign-restricted-workflows.html).
The [reference index](https://davidzenz.github.io/bsvarPost/reference/index.html) lists lower-level variants and
plotting methods without repeating them here.
