---
title: "Post-Estimation Workflows in bsvarPost"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Post-Estimation Workflows in bsvarPost}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

This vignette focuses on the broader analytical surface of `bsvarPost` after a
model has already been estimated with `bsvars` or `bsvarSIGNs`.

The emphasis is not on estimation itself, but on:

- representative-model summaries
- posterior hypothesis and restriction auditing
- event-window historical decomposition workflows
- response timing summaries
- diagnostics for sign-restricted models
- plot styling and publication-oriented outputs

## Representative-model summaries

For sign-restricted analysis, pointwise posterior medians need not correspond to
one coherent draw. `bsvarPost` therefore offers representative-model summaries.

```{r}
library(bsvarSIGNs)
library(bsvarPost)

data(optimism)

sign_irf <- matrix(c(0, 1, rep(NA, 23)), 5, 5)

spec <- specify_bsvarSIGN$new(
  optimism * 100,
  p = 4,
  sign_irf = sign_irf
)

post <- estimate(spec, S = 100, thin = 1, show_progress = FALSE)

rep_median <- median_target_irf(post, horizon = 12)
rep_mla <- most_likely_admissible_irf(post, horizon = 12)

summary(rep_median)
summary(rep_mla)
plot(rep_median)
```

Use `median_target_*()` when you want the stored draw closest to the posterior
center. Use `most_likely_admissible_*()` when you want the stored admissible
draw with the highest reconstructed admissibility score.

## Posterior hypotheses and audits

Posterior probability statements can be evaluated directly on IRFs or CDMs.

```{r}
hypothesis_irf(
  post,
  variable = 1,
  shock = 1,
  horizon = 0:2,
  relation = ">",
  value = 0
)

joint_hypothesis_irf(
  post,
  variable = 1,
  shock = 1,
  horizon = 0:2,
  relation = ">",
  value = 0
)

simultaneous_irf(
  post,
  horizon = 12,
  variable = 1,
  shock = 1
)
```

Restriction and magnitude audits are explicit post-estimation summaries. They
do not modify the estimation problem.

```{r}
restriction_audit(post)

magnitude_audit(
  post,
  type = "irf",
  variable = 2,
  shock = 1,
  horizon = 0,
  relation = ">",
  value = 0
)
```

## Historical decomposition workflows

For event studies, aggregate historical decomposition contributions over a
window and then rank shocks by contribution size.

```{r}
hd_event <- tidy_hd_event(post, start = 1, end = 4)
hd_event

ranked <- shock_ranking(post, start = 1, end = 4, ranking = "absolute")
ranked
```

Compare event windows across models:

```{r}
compare_hd_event(p1 = post, p2 = post, start = 1, end = 4)
```

Plot the resulting summaries:

```{r}
plot_hd_event(post, start = 1, end = 4)
plot_shock_ranking(post, start = 1, end = 4, ranking = "absolute", top_n = 5)
```

## Response timing summaries

`bsvarPost` includes timing summaries that applied users often compute manually
for papers and presentations.

```{r}
peak_response(post, type = "irf", horizon = 12, variable = 1, shock = 1)

duration_response(
  post,
  type = "cdm",
  horizon = 12,
  variable = 1,
  shock = 1,
  relation = ">",
  value = 0,
  mode = "total"
)

half_life_response(
  post,
  type = "irf",
  horizon = 12,
  variable = 1,
  shock = 1,
  baseline = "peak"
)

time_to_threshold(
  post,
  type = "cdm",
  horizon = 12,
  variable = 1,
  shock = 1,
  relation = ">",
  value = 0
)
```

The same summaries can be compared across alternative specifications.

```{r}
compare_peak_response(p1 = post, p2 = post, type = "irf", horizon = 12, variable = 1, shock = 1)
compare_duration_response(p1 = post, p2 = post, type = "cdm", horizon = 12, variable = 1, shock = 1, relation = ">", value = 0)
compare_half_life_response(p1 = post, p2 = post, type = "irf", horizon = 12, variable = 1, shock = 1)
compare_time_to_threshold(p1 = post, p2 = post, type = "cdm", horizon = 12, variable = 1, shock = 1, relation = ">", value = 0)
```

## Diagnostics for sign-restricted models

`acceptance_diagnostics()` provides a stored-draw diagnostics layer for
`PosteriorBSVARSIGN` objects.

```{r}
diag_tbl <- acceptance_diagnostics(post)
diag_tbl

summary(diag_tbl)
```

Compare diagnostics across models:

```{r}
compare_acceptance_diagnostics(p1 = post, p2 = post)
plot_acceptance_diagnostics(diag_tbl, metrics = c("effective_sample_size", "kernel_zero_share"))
```

These are not full proposal acceptance histories. They are summaries of what is
recoverable from the stored posterior draws and identification state.

## Plot styling and templates

All `ggplot2` outputs returned by `bsvarPost` can be restyled downstream.

```{r}
irf_tbl <- tidy_irf(post, horizon = 12)

p <- ggplot2::autoplot(irf_tbl)

style_bsvar_plot(
  p,
  preset = "paper",
  palette = c("#1b9e77", "#d95f02")
)

template_bsvar_plot(
  p,
  family = "irf",
  preset = "paper"
)

annotate_bsvar_plot(
  p,
  title = "Impulse responses",
  xintercept = 2
)
```

This plotting layer is intended to reduce the amount of manual figure code
needed to bring posterior outputs into publication-ready form.
